---
title: "Visualizing Distribution 1"
output: github_document
---

```{r global_options, include=FALSE}
library(knitr)
library(ggplot2)
opts_chunk$set(fig.align="center", fig.height=4, fig.width=5.5)
```

In this notebook I will be working with the `chickwts` dataset built into R. This data set contains weight measurements of chicks fed on different food sources to assess their effectiveness on growth rate.

```{r}
head(chickwts)
```

**Viz 1:** Use ggplot to make a histogram of the `weight` column. Manually choose appropriate values for `binwidth` and `center`. Explain your choice of values in 2-3 sentences.

```{r}
ggplot(chickwts, aes(weight)) +
  geom_histogram(binwidth = 20, center = 110, color= 'black', fill = 'skyblue') +
  scale_x_continuous( breaks= seq(from = 110, to = 450, by = 40) ) + 
  scale_y_continuous( breaks= seq(from = 0, to = 14, by = 2) )

```

Range of weights is from 108 to 423 lbs. The difference between the 5th percentile (140.5) and the 95th percentile (385.0) is
244.5 lbs. The sample size is 71. A bindiwth of 10 is spiky and doesn't seem representative as it shows many gaps in intermediate weights.  A bindwith of 20 lbs seems to be a reasonable quantity when talking about weights and it is also 7% of the mean value.

**Viz 2:** Modify the plot from Problem 1 to show one panel per feed. Hint: Use `facet_wrap()`.

```{r}
ggplot(chickwts, aes(weight, fill = feed)) +
  geom_histogram(binwidth = 20, color = "black", center = 110) + facet_wrap(vars(feed)) + 
  scale_x_continuous( breaks= seq(from = 100, to = 450, by = 75) )
```

