---
title: "Food choices"
---

```{r setup, include=FALSE}
library(tidyverse)
library(colorspace)
knitr::opts_chunk$set(echo = TRUE)
```

This is the dataset we will be working with:

```{r message = FALSE, warning = FALSE}
food <- readr::read_csv(
  "https://wilkelab.org/DSC385/datasets/food_coded.csv",
  na = c("Unknown", "nan", "Personal"))
food
```

A detailed data dictionary for this dataset is available [here.](https://wilkelab.org/DSC385/datasets/food_codebook.pdf) The dataset was originally downloaded from Kaggle, and you can find additional information about the dataset [here.](https://www.kaggle.com/borapajo/food-choices/version/5)


## **Question:** 
Is GPA related to student income, the father's educational level, or the student's perception of what an ideal diet is?


## **Introduction:**


## **Approach:** 


## **Analysis:**


### 1.0 Data Inspection
```{r }
# Get only the four columns of interest
food <- food[c("GPA","income","father_education", "ideal_diet_coded")]
colnames(food)
```
####1.1 `GPA`

Missing values encoded as "Unknown", "nan", or "Personal" where parsed as valid NAs at reading time. However, one more entry has an invalid string, which is handled below.

```{r }
food$GPA <- sub(' bitch','',food$GPA)
```

Convert to numeric
```{r }
food$GPA <- as.numeric(food$GPA )
```

```{r }
summary(food$GPA)
```


```{r}
str_detect(food$GPA, 'bitch')
```
```{r}
str_detect(food$GPA, 'bitch')
```



```{r }

summary(food$income)
sum(is.na(food$income))

```
```{r }

summary(food$father_education)
sum(is.na(food$father_education))

```

```{r }

summary(food$ideal_diet_coded)
sum(is.na(food$ideal_diet_coded))

```
## **Discussion:**


