---
title: "Intro to ggplot"
author: "By Lorenzo Peve"
output:
  github_document:
    toc: true

---

## 1.0 Load Libraries
```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(knitr)
library(ggplot2)
library(gapminder)
library(ggthemes)
```

## 2.0 View at *gapminder* data

```{r}
gapminder
```

## 3.0 Plotting Data

Different plots are going to be created. I'll start with the most basic plot and
then be be gradually increasing the number of components and refining the graph
to make the most effective visualization possible.

### 3.1 Most basic plot
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp)) + geom_point()
p
```

### 3.2 Adding color
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp, color = continent)) + geom_point()
p
```

### 3.3 Smoothing
`geom_smooth()` has calculated a smoothed line and added a shaded ribbon showing
the standard error for the line.
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y=lifeExp))
p + geom_smooth()
```

### 3.4 Smoothing with points 
The order of the layers matter. We want points below the smoothing, so they go
first.
```{r}
p + geom_point(color = '#006600') + geom_smooth()
```

### 3.5 Transforming x-axes.

GDP is not normally distributed across the country years. The x-axis scale would
probably look better if it were transformed from a linear scale to a log scale. 

```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y=lifeExp))
p + geom_point() +
    geom_smooth(method = "lm") +
    scale_x_log10()
```

### 3.6 Styling Axes

GDP is not normally distributed across the country years. The x-axis scale would
probably look better if it were transformed from a linear scale to a log scale. 

```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y=lifeExp))
p + geom_point(color = 'firebrick') +
    geom_smooth(method = "lm") +
    scale_x_log10(name = 'GDP (U$D)', 
                  labels = scales::dollar, 
                  breaks = waiver(),
                  minor_breaks= waiver()) + 
    scale_y_continuous(name = 'Life Expectancy',
                       limits = c(25, 90),
                       n.breaks = 6) +
    theme_tufte()
```

### 3.7 A more stylish plot of Life Expectancy vs GDP

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y=lifeExp))
p + geom_point(alpha = 0.3) +
    geom_smooth(method = "lm") +
    scale_x_log10(labels = scales::dollar) +
    labs(x = "GDP Per Capita", y = "Life Expectancy in Years",
         title = "Economic Growth and Life Expectancy",
         subtitle = "Data points are country-years",
         caption = "Source: Gapminder.")
```

### 3.8 Mapping the continent variable

```{r}
p <- ggplot(data = gapminder, 
            mapping = aes(x = gdpPercap, 
                          y=lifeExp, 
                          color = continent,
                          fill = continent))
p + geom_point(alpha = 0.4) +
    geom_smooth() +
    scale_x_log10(labels = scales::dollar) +
    labs(x = "GDP Per Capita", y = "Life Expectancy in Years",
         title = "Economic Growth and Life Expectancy",
         subtitle = "Data points are country-years",
         caption = "Source: Gapminder.")
```

### 3.9 Mapping Aestheticsper geom

Five separate smoothers is too many but we still want to show the smoothed line.
By default, **geoms** inherit their mappings from the `ggplot()` function. 

```{r}
p <- ggplot(data = gapminder, 
            mapping = aes(x = gdpPercap, 
                          y=lifeExp))
p + geom_point(mapping = aes(color = continent), alpha = 0.4) +
    geom_smooth() +
    scale_x_log10(labels = scales::dollar) +
    labs(x = "GDP Per Capita", y = "Life Expectancy in Years",
         title = "Economic Growth and Life Expectancy",
         subtitle = "Data points are country-years",
         caption = "Source: Gapminder.")
```

### 3.10 Color mapping population

```{r}
p <- ggplot(data = gapminder, 
            mapping = aes(x = gdpPercap, 
                          y=lifeExp))
p + geom_point(mapping = aes(color = log(pop)), alpha = 1) +
    geom_smooth(color = 'red') +
    scale_x_log10(labels = scales::dollar) +
    labs(x = "GDP Per Capita", y = "Life Expectancy in Years",
         title = "Economic Growth and Life Expectancy",
         subtitle = "Data points are country-years",
         caption = "Source: Gapminder.")
```