---
title: "Project 1"
output: github_document
---

```{r global_options, include=FALSE}
library(knitr)
library(ggplot2)
opts_chunk$set(fig.align="center", fig.height=4, fig.width=5.5)
```

**Part 1:** Demonstrate basic command of Markdown by creating a bulleted list with three items, a numbered list with three items, and a sentence that has one word in bold and one word in italics.

A few distinct characteristics of `ggplot2` compared to other libraries is that
`ggplot2` :

* Has an underlying *grammar* based on the **Grammar of Graphics** that allows
you to compose graphs by combining independent components;
* Works iteratively by starting with the raw data and then a sequence of logical
steps that add to the structure of the plot; 
* Has carefully chosen default values that can produce publication-quality
graphics with a single function.

The process of making plots using `ggplot()` can be explained by three distinct
steps:

1. Provide the data and specify the mapping
    + 1.1 Data is generally provided as a dataframe
    + 1.2 The mapping is done with the function `aes()` that connects variables
    in the dataframe to aesthetics (position in x or y, color, shape, size, etc)
2. Specify which type of plot you want. How do you want `ggplot()` to draw
the data.
    + Functions like `geom_point()` or `geom_line()` are used in this step.
    + Arguments can be passed to the `geom_`
3. Use of additional functions can be used to adjust scales, labels, tick marks,
and or titles.

It is noted that with only Steps 1 and 2 there is enough information for
`ggplot()` to draw a **plot**. If you don’t provide anything else `ggplot()`
will use *carefully* designed defaults.


**Part 2:** The `mtcars` dataset contains information on car design and performance in 1974:

```{r}
show(head(mtcars))
```

Use ggplot to make a line plot of the fuel economy of a car, measured in miles/gallon (column `mpg`), versus its weight (column `wt`).

```{r}
p <- ggplot(data = mtcars, 
            mapping = aes(x = wt, 
                          y=mpg))
p + geom_line() + 
scale_y_continuous(limits = c(5, 35), breaks = seq(5, 35, by=5)) + 
scale_x_continuous(limits = c(1, 5.5), breaks = seq(1, 5.5, by=1)) +
labs(x = "Weight (1000 lbs)", y = "Miles/(US) gallon",
     title = "Fuel Economy of Cars",
     caption = "Source: 1974 Motor Trend US magazine.")

```

Now make a scatter plot (using `geom_point()`) of the horse power of a car (`hp`) versus its weight, and color points by its fuel economy.

```{r}
p <- ggplot(data = mtcars,
            mapping = aes(x = wt,
                          y=hp,
                          color = mpg))
p + geom_point(size = 4) +
scale_y_continuous(limits = c(50, 350), breaks = seq(50, 350, by=50)) +
scale_x_continuous(limits = c(1, 5.5), breaks = seq(1, 5.5, by=1)) +
labs(x = "Weight (1000 lbs)", y = "Gross horsepower",
     title = "Horse Power of Cars",
     caption = "Source: 1974 Motor Trend US magazine.")
```
