---
title: "Project 2"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

This is the dataset for the project:
```{r message = FALSE}
olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')

olympic_gymnasts <- olympics %>% 
  filter(!is.na(age)) %>%             # only keep athletes with known age
  filter(sport == "Gymnastics") %>%   # keep only gymnasts
  mutate(
    medalist = case_when(             # add column for success in medaling
      is.na(medal) ~ FALSE,           # NA values go to FALSE
      !is.na(medal) ~ TRUE            # non-NA values (Gold, Silver, Bronze) go to TRUE
    )
  )
olympic_gymnasts

```

More information about the dataset can be found at https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-07-27/readme.md and https://www.sports-reference.com/olympics.html.

## **Question:** 
Are there age differences for male and female Olympic gymnasts who were successful or not in earning a medal, and how has the age distribution changed over the years? 


## **Introduction:** 
This project uses a subset of the historical dataset on the modern **Olympic Games** (including all the Games from Athens 1896 to Rio 2016) by focusing on gymnasts. `olympic_gymnasts` contains all the records from the original **Olympic Games** dataset where the sport is equal to "Gymnastics". In this dataset, each row corresponds to the participation of an Olympic athlete in an Olympic event. 

It is noted that an athlete is allowed to participate in multiple events during the same Olympic Games. Thus, the same athlete (id, name, sex, age, height, team, noc) is repeated for those athletes that took part in different competitions during the same Games. 

It is also important to highlight that the athlete data (age, height, weight) correspond to the time of the competition. Athletes that took part in more than one Game, have larger ages for more recent Games (i.e., naturally aging).

There are 16 columns providing information about the athlete and their participation/performance. Information about the athlete includes:

* id, sex, age, height, weight, team, and nationality.

Information about the athlete participation/performance includes:

* game's name, year, season, city
* athlete's sport/discipline, event, medal
* whether the athlete was a medalist.


## **Approach:** 
The approach to evaluating differences between male and female Olympic gymnasts while distinguishing those with medals and those without consisted of using violin plots (`geom_violin()`). 

To evaluate how has the age distribution changed over the years, a facetted boxplot is used.

## **Analysis:** 


### 1.0 Data Validation. 
Check there are no missing values.
```{r }
unique(olympic_gymnasts$sex)
unique(olympic_gymnasts$medalist)
```
Get range of ages

```{r }
summary(olympic_gymnasts$age)
```

### 2.0 Violin Plots


```{r }

ggplot(olympic_gymnasts, aes(x = medalist, y = age, fill = sex)) +
geom_violin() + 
scale_x_discrete(name = NULL,
                 labels = c("No Medal", "Medalled")) +
scale_y_continuous(breaks= seq(from = 10, to = 50, by = 5),
                  name = 'Athlete Age during Competition') +
scale_fill_manual(name = 'Gender',
                  labels = c("Male", "Female"),
                  values = c(`M` = "#3578FD", `F` = "#E69F00")
  ) +
theme_bw(12)

```

### 3.0 Box Plots

```{r}

ggplot(olympic_gymnasts, aes(x = year, y = age, group = year)) +
geom_boxplot() + 

facet_wrap(vars(medalist), 
           labeller = as_labeller(c(`TRUE` = "medaled", `FALSE` = "did not medal")) ) +
scale_x_continuous(breaks= seq(from = 1890, to = 2030, by = 20),
                  name = 'Athlete Age during Competition')

```

## **Discussion:** 

Interestingly, gymnast's "medalling" status does not affect the age distribution within a particular gender. We can see this by comparing the violins with the same color and seeing that each pair has a similar shape and expected values. 
Hence, comparisons are made for gender only. For both medalled and not medalled athletes, the expected value for the age of an Olympic women athlete is considerably smaller than that of males, with women on average being eight years younger than male competitors.
Besides, looking at the shape of the distribution, it seems that the age of male participants resembles more closely a bell-shape curved. In contrast, the distribution of age in female participants is more right-skewed, with a bigger proportion of younger athletes.

The distribution of the age of Olympic gymnasts doesn't exhibit a difference in the trend when distinguished by medalling status. As seen in both subplots, the trend is for ages to decrease since the 1960's and remain about 20 years old for the last 30 years. However, the distribution of ages for non-medallist athletes is more spread, as indicated by the length of the boxplot's fence lines.

